---
title: "Perceptual Maps and Customer Segments"
subtitle: "TBFI CEP Project: My AO2 Proposal"
author: "Simran Kaur"
institute: "MS Digital Marketing & Analytics, Cal Poly Pomona"
date: today
format:
  revealjs:
    theme: serif
    include-in-header:
      text: |
        <style>
        :root { --r-background-color: #fbf8f3; --r-main-color: #26342f; --r-heading-color: #234c43; --r-link-color: #816038; --r-main-font: Georgia, serif; --r-heading-font: Georgia, serif; --r-main-font-size: 38px; }
        .reveal h1 { font-size: 1.8em; }
        .reveal h2 { font-size: 1.4em; margin-bottom: .6em; }
        .reveal .slides { text-align: left; }
        .reveal .slides section#title-slide { text-align: center; }
        .reveal table { font-size: .86em; width: 100%; }
        .reveal table th { color: #234c43; border-bottom: 2px solid #234c43; }
        .reveal table td { padding: .35em .45em; }
        .reveal .footer { font-size: 18px; }
        .reveal .aside { font-size: .65em; color: #6c655d; }
        .reveal blockquote { background: transparent; box-shadow: none; border-left: 4px solid #b39164; padding-left: 1em; }
        </style>
    html-math-method: plain
    width: 1600
    height: 900
    footer: "TBFI CEP · Simran Kaur"
    transition: slide
    incremental: false
    toc: false
    slide-number: true
    scrollable: true
    smaller: true
    code-overflow: wrap
    embed-resources: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
  freeze: auto
---

```{r}
#| label: setup
#| include: false
# Invented practice data, not actual CEP results.
demo_ratings <- data.frame(
  respondent_id = rep(sprintf("D%02d", 1:4), each = 3),
  brand = rep(c("Brand A", "Brand B", "Brand C"), times = 4),
  affordability = c(5, 3, 2, 4, 3, 1, 5, 4, 2, 4, 2, 1),
  quality = c(2, 4, 5, 2, 4, 4, 3, 5, 5, 1, 3, 4)
)
```

## Contents

-   Project introduction
-   My objective and why it matters
-   Literature review
-   Data and methods
-   Reproducible example
-   References and project links

## Project Introduction

Our TBFI CEP project explores an **AI consultant-training simulation and an industry insights dashboard**.

-   We plan to compare AI survey responses with a human survey benchmark.
-   The project focuses on four industries in the **San Gabriel Valley (SGV)**.
-   The business problem is understanding how customers see brands and what different customers need.

::: notes
This is a proposal. I am explaining what we plan to do, rather than presenting final survey findings.
:::

## My Objective: AO2

> How do customers see brands, and what customer groups can we identify?

::::: columns
::: {.column width="48%"}
### Perceptual Maps

Show which brands customers see as similar or different.
:::

::: {.column width="48%"}
### Customer Segments

Group customers who have similar preferences or needs.
:::
:::::

**Why it matters:** These results can help business owners decide how to position their brands and whom to target.

::: notes
The map is about brands. Customer segments are about people. My role is to make both easier to understand visually.
:::

## Literature Review

**Smith (1956): Market segmentation**

This marketing framework helps explain why businesses should consider differences in customer needs.

**Kruskal (1964): Multidimensional scaling**

MDS is a method for showing similarities and differences as distances on a map.

**What we still need to check:** Whether our survey includes useful brand ratings and whether AI responses reflect patterns in the human data.

::: aside
Full references appear on the final slide.
:::

## Data and Sampling

| Data source | What we need to know |
|:-----------------------------------|:-----------------------------------|
| Human survey | How participants were recruited and the final respondent count |
| AI responses | Which model and prompt generated each response |
| Both sources | Whether the questions and rating scales are comparable |

**Sample characteristics:** We will describe respondent counts and available demographics after reviewing the data.

::: notes
The final sample size and industry names need confirmation. We will distinguish generated AI records from real survey participants.
:::

## Data Preparation and Measures

::: incremental
-   Check respondent IDs and accidental duplicate records.
-   Review missing answers and values outside the rating scale.
-   Keep ratings numeric and ZIP codes as text.
-   Choose relevant measures for each industry, such as affordability or quality.
:::

**Goal:** Prepare consistent data before creating maps or customer groups.

::: notes
A person may rate several brands, so a repeated ID is not always an error. We need to understand what each row represents.
:::

## Planned Analysis

**Perceptual maps:** We plan to use **multidimensional scaling (MDS)** to show how customers see brands.

**Customer segments:** We plan to use **clustering** to group customers with similar preferences.

**Comparison:** We will examine similarities and differences between AI and human results.

::: notes
MDS creates a picture of brand relationships. Clustering identifies groups of similar customers. Our exact choices will depend on the available data.
:::

## Reproducible Example

**Practice data:** `r length(unique(demo_ratings$respondent_id))` invented respondents each rate `r length(unique(demo_ratings$brand))` brands on a 1–5 scale.

```{r}
#| label: tbl-demo-ratings
#| tbl-cap: "Average practice ratings by brand"
# Calculate average ratings, then display them.
brand_means <- aggregate(
  cbind(affordability, quality) ~ brand,
  data = demo_ratings,
  FUN = mean
)
knitr::kable(
  brand_means,
  col.names = c("Brand", "Affordability", "Quality"),
  digits = 1
)
```

In this example, Brand A has the highest affordability rating, while Brand C has the highest quality rating.

::: aside
Higher affordability means more affordable. These invented ratings demonstrate the R workflow, not project findings or an MDS map.
:::

::: notes
R calculates this table when I render the slides. This meets the reproducible-table requirement and shows a simple way to summarize brand ratings.
:::

## References and Appendix {.smaller}

Smith, W. R. (1956). Product differentiation and market segmentation as alternative marketing strategies. *Journal of Marketing, 21*(1), 3–8. [Read source](https://doi.org/10.1177/002224295602100102)

Kruskal, J. B. (1964). Multidimensional scaling by optimizing goodness of fit to a nonmetric hypothesis. *Psychometrika, 29*, 1–27. [Read source](https://doi.org/10.1007/BF02289565)

### Appendix: Clickable Links

-   [Group 8 CEP report](https://gabrielarq1.github.io/CEP--Group-8/CEP-F26/Group8-CEP-F26.html)

-   [Group 8 repository](https://github.com/gabrielarq1/CEP--Group-8)

-   [My M06 presentation files](https://github.com/Simrank-DM/IBM6500-F26/tree/main/MO6)